### 2017-11-27 paper/sections/1_introduction.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different domain and service.  Our exploration of this space results in a
dynamically sized cache policy that sacrifices negligible performance while
using 32-66\% less memory than the default ParSplice configuration.

@@

%very customized solution confirms that our prototype can adapt to different
%types of workloads; but more importantly, this is the second system and domain
%that we have successfully used this policy language and engine in. This breadth
%helps us reason about more general data management policies by giving us a
%framework to compare and contrast policies from other domains and services.

@@

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
was designed for the narrow purpose of file system metadata load balancing,
this paper provides evidence that the approach is more broad.  Specifically,
the collection of abstractions in Mantle provide a general control plane that
improves the performance of metadata access. So in this paper we refer to
Mantle as a policy engine that injects policies written in our data management
language directly into a running service, such as a file system or key-value
store.  Rather than co-designing a policy directly into the service, which
requires (1) domain-specific knowledge about the application to design an
effective policy and (2) service-specific knowledge to find and change
policies, we separate policies from mechanisms in a general way. By exposing
these policies (and not the mechanism) developers unfamiliar with the service
that their application uses can quickly deploy solutions based on their
domain-specific knowledge.  We show that our framework:

>>>>

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different application and storage system.  Our exploration of this space results in a
dynamically sized cache policy that sacrifices negligible performance while
using 32-66\% less memory than the default ParSplice configuration.

@@

@@

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  While Mantle
was designed for the narrow purpose of file system metadata load balancing,
this paper provides evidence that the approach is more broad.  Specifically,
the collection of abstractions in Mantle provide a general control plane that
improves the performance of metadata access. So in this paper we refer to
Mantle as a policy engine that injects policies written in our data management
language directly into a running storage system, such as a file system or key-value
store.  Rather than implementing a policy directly into the storage system,
which requires domain-specific knowledge to find and change policies in the
storage system, we separate policies from mechanisms in a general way.  By
exposing just the policies developers unfamiliar with the storage system that
their application uses can quickly deploy efficient solutions based on their
application-specific knowledge.  We show that our framework:

### 2017-11-27 paper/sections/3_methodology.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

%  \item motivation: Mochi load balancer microservice
%  \item library architecture, callbacks, environment

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a service.  Administrators write policies in Lua from whatever
domain they choose ({\it e.g.}, statistics, machine learning, storage system)
and the policies are embedded into the runtime by Mantle.  We chose Lua for
simplicity, performance, and portability; it is a scripting language with
simple syntax, which allows administrators to focus on the policies themselves;
it was designed as an embeddable language, so it is lightweight and does less
type checking; and it interfaces nicely with C/C++.  When the service makes
decisions it executes the administrator-defined policies for when/where/how
much and returns a decision.  To do this, the service needs to be modified to
(1) provide an environment of metrics and (2) identify where policies are set.
These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

@@

\begin{table}
  \centering
  \begin{tabular}{ r | l | l }
  Metrics     & Data Structure & Description \\\hline
  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
              & resource util. for servers \\
  Time Series & [(ts, val), ..., (ts, val)]
              & accesses by timestamp (ts) \\
  && \\
              & Service      & Example \\\hline
  Cluster     & File Systems & CPU util., Inode reads \\
              & ParSplice    & CPU util., Cache size \\
  Time Series & File Systems & Accesses to directory \\
              & ParSplice    & Accesses to key in DB\\
  \end{tabular}
  \caption{Types of metrics exposed by the service to the policy engine using Mantle.\label{table:metrics}}
\end{table}

\subsubsection{Environment of Metrics} services expose \textbf{cluster} metrics
for describing resource utilization and \textbf{time series} metrics for
describing accesses to some data structure over time. Table~\ref{table:metrics}
shows how these metrics are accessed from the policies written by
administrators.

For cluster metrics, the service passes a dictionary to Mantle. Policies access
the cluster metric values by indexing into a Lua table using \texttt{server}
and \texttt{metric}, where server is a node identifier ({\it e.g.}, MPI Rank,
metadata server name) and metric is a resource name.  Metrics used for file
system metadata load balancing are shown by the ``environment" box in
Figure~\ref{fig:mantle}. The measurements and exchange of metrics between
servers is done by the service; Mantle in CephFS leverages metrics from other
servers collected using CephFS's heartbeats.  For example, a policy written for
an MPI-based service can access the CPU utilization of the first rank in a
communication group using:

@@

For time series metrics, the service passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses.  The
service uses a pointer to the time series to facilitate time series with many
values, like accesses to a database or directory in the file system namespace.
This decision limits the time series metrics to only include values from the
{\it current } node, although this is not a limitation of Mantle itself.  For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
d = timeseries()       -- get data (d) from service
for i=1,d:size() do    -- iterate over timeseries
  ts, value = d:get(i) -- index into timeseries 
  if value == 'mydirectory' then
    count = count + 1
  end
end
\end{minted}
\end{figure}

>>>>

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a storage system.  Administrators write policies in Lua from whatever
domain they choose ({\it e.g.}, statistics, machine learning, storage system)
and the policies are embedded into the runtime by Mantle.  We chose Lua for
simplicity, performance, and portability; it is a scripting language with
simple syntax, which allows administrators to focus on the policies themselves;
it was designed as an embeddable language, so it is lightweight and does less
type checking; and it interfaces nicely with C/C++.  When the storage system makes
decisions it executes the administrator-defined policies for when/where/how
much and returns a decision.  To do this, the storage system needs to be modified to
(1) provide an environment of metrics and (2) identify where policies are set.
These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

@@

\begin{table}
  \centering
  \begin{tabular}{ r | l | l }
  Metrics     & Data Structure & Description \\\hline
  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
              & resource util. for servers \\
  Time Series & [(ts, val), ..., (ts, val)]
              & accesses by timestamp (ts) \\
  && \\
              & Storage System & Example \\\hline
  Cluster     & File Systems & CPU util., Inode reads \\
              & ParSplice    & CPU util., Cache size \\
  Time Series & File Systems & Accesses to directory \\
              & ParSplice    & Accesses to key in DB\\
  \end{tabular}
  \caption{Types of metrics exposed by the storage system to the policy engine using Mantle.\label{table:metrics}}
\end{table}

\subsubsection{Environment of Metrics} storage systems expose \textbf{cluster} metrics
for describing resource utilization and \textbf{time series} metrics for
describing accesses to some data structure over time. Table~\ref{table:metrics}
shows how these metrics are accessed from the policies written by
administrators.

For cluster metrics, the storage system passes a dictionary to Mantle. Policies access
the cluster metric values by indexing into a Lua table using \texttt{server}
and \texttt{metric}, where server is a node identifier ({\it e.g.}, MPI Rank,
metadata server name) and metric is a resource name.  Metrics used for file
system metadata load balancing are shown by the ``environment" box in
Figure~\ref{fig:mantle}. The measurements and exchange of metrics between
servers is done by the storage system; Mantle in CephFS leverages metrics from other
servers collected using CephFS's heartbeats.  For example, a policy written for
an MPI-based storage system can access the CPU utilization of the first rank in a
communication group using:

@@

For time series metrics, the storage system passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses.  The
storage system uses a pointer to the time series to facilitate time series with many
values, like accesses to a database or directory in the file system namespace.
This decision limits the time series metrics to only include values from the
{\it current } node, although this is not a limitation of Mantle itself.  For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
d = timeseries()       -- data (d) from storage system
for i=1,d:size() do    -- iterate over timeseries
  ts, value = d:get(i) -- index into timeseries 
  if value == 'mydirectory' then
    count = count + 1
  end
end
\end{minted}
\end{figure}

### 2017-11-27 paper/sections/4_dynamic-policies.tex
paper: addresses #47
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Cache Management Using System Architecture Knowledge}
\label{sec:arch-specific}

@@

% technical details
The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The middle graph labeled "Fixed Cache Size" shares the \(y\) axis and shows the
trade-off of using a basic LRU-style cache for different cache sizes, where the
penalty for a cache miss is retrieving the data from the persistent database.
We evict keys (if necessary) at every operation instead of when segments
complete because the cache fills up too quickly otherwise. This is implemented
using a ``when" policy of:

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs.  wall-clock time curves down over time; as the nanoparticle grows it
takes longer to generate segments so by the time we reach 2 hours, over 90\% of
the trajectory is already generated.  For (2), the memory footprint rises until
we reach 100K key switch threshold at 0.4GB, as shown by the ``\(\Delta_1\),
Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
conclude that this policy only works well for this 2.5 hour run.

% but the result is still valid
Despite these caveats, the result is still valid: we found a multi-policy cache 
management strategy that absorbs the cost of a high read throughput on a small
keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
show the effectiveness of the policy engine we integrated into
ParSplice, not that we were able to identify the best policy for all system
setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
job lengths).  To solve that problem, we need a way to identify what thresholds
we should use for different configurations.

\section{Cache Management Using Domain-Specific Application Knowledge}
\label{sec:dom-specific}

\begin{figure*}[t!]
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
\caption{Key activity for a 4 hour run shows groups of accesses to the same
subset of keys. Detecting these access patterns leads to a more accurate cache
management strategy, which is discussed in
Section~\S\ref{sec:regime-detection}.\label{fig:keyspace-zoomed}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.5\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/dscache-vs-none.png}\\
	\caption{Performance/Utilization for dynamically sized cache (DSCache)
policy. With negligible performance degradation, DSCache adjusts to different
\(\Delta\)s over 2\(\times\) memory in the best case.
\label{fig:dscache-vs-none}}
    \end{subfigure}%
    \caption{Different cache management policies tested over the Mantle policy engine.}
\end{figure*}

@@

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.  The goal of the following
sections is not to find an optimal solution, as this can be done with parameter
sweeps for thresholds; rather, we try to find techniques that work for a range
of inputs and system setups.

>>>>

\section{Cache Management Using Storage System Architecture Knowledge}
\label{sec:arch-specific}

@@

% technical details
The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The middle graph labeled ``Fixed Cache Size" shares the \(y\) axis and shows the
trade-off of using a basic LRU-style cache for different cache sizes, where the
penalty for a cache miss is retrieving the data from the persistent database.
We evict keys (if necessary) at every operation instead of when segments
complete because the cache fills up too quickly otherwise. This is implemented
using a ``when" policy of:

@@

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs.  wall-clock time curves down over time; as the nanoparticle grows it
takes longer to generate segments so by the time we reach 2 hours, over 90\% of
the trajectory is already generated.  For (2), the memory footprint rises until
it reaches the 100K key switch threshold at 0.4GB and then reduces to the final
value after switching policies. The memory usage over time for this policy is
shown by the ``\(\Delta_1\), Multi-Policy" curve in
Figure~\ref{fig:memory-vs-time} but in Figure~\ref{fig:methodology-tradeoff} we
plot the final value.  Despite these caveats, the result is still valid: we
found a multi-policy cache management strategy that absorbs the cost of a high
read throughput on a small keyspace and reduces the memory pressure for a 2.5
hour run. Our experiments show the effectiveness of the policy engine we
integrated into ParSplice, not that we were able to identify the best policy
for all system setups ({\it i.e.} different ParSplice parameters, number of
worker tasks, and job lengths).  To solve that problem, we need a way to
identify what thresholds we should use for different configurations.

\section{Cache Management Using Application-Specific Knowledge}
\label{sec:dom-specific}

\begin{figure*}[t!]
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
	\includegraphics[width=1\textwidth]{figures/keyspace-zoomed.png}\\
\caption{Key activity for a 4 hour run shows groups of accesses to the same
subset of keys. Detecting these access patterns leads to a more accurate cache
management strategy, which is discussed in
Section~\S\ref{sec:regime-detection}.\label{fig:keyspace-zoomed}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/dscache-vs-none.png}\\
	\caption{Performance/Utilization for dynamically sized cache (DSCache)
policy. With negligible performance degradation, DSCache adjusts to different
\(\Delta\)s over 2\(\times\) memory in the best case.
\label{fig:dscache-vs-none}}
    \end{subfigure}%
    \caption{Different cache management policies tested over the Mantle policy engine.}
\end{figure*}

@@

Feeding application-specific knowledge about the ParSplice application into a policy
leads to a more accurate cache management strategy.  The goal of the following
sections is not to find an optimal solution, as this can be done with parameter
sweeps for thresholds; rather, we try to find techniques that work for a range
of inputs and system setups.

### 2017-11-27 paper/sections/5_file-systems.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

In the previous section, we used our data management language and the Mantle
policy engine to design effective cache management strategies for a new service
and domain. In this section, we compare and contrast the policies examined for
file system metadata load balancing in~\cite{sevilla:sc15-mantle} with the ones
we designed and evaluated above for cache management in ParSplice.

@@

The cache management policies we developed earlier can be used by load
balancing policies to effectively spread load across a cluster. For example,
many distributed file systems spread the hierarchical namespace across a
dedicated metadata cluster and caching policies could be used to determine what
metadata to move and when to move it. To demonstrate this idea, consider
Figure~\ref{fig:trace-atime}, where the dots are the file system metadata reads
in a 1 hour window of a 3-day Lustre namespace trace and the vertical lines are
the groups of accesses detected by our policy from
Section~\S\ref{sec:regime-detection}.  The \(x\) axis is time and the \(y\)
axis is the file ID, where the file IDs are listed in the order that they
appear in the trace.

The trace is anonymized so all file names are replaced with a unique identifier
and we do not know what clients are doing.  Although other access pattern
detection algorithms are possible, we use the one designed for cache management
in Section~\S\ref{sec:regime-detection} with slight modifications based on our
knowledge of file systems~\footnote{We filtered requests key ids less than
2000, as these are most likely path traversal requests to higher parts of the
namespace, {\it i.e.} root directories.} and it successfully detects the
largest group of key accesses that starts at time 1000 seconds and ends at time
2200 seconds. File systems that load balance file system metadata across a
cluster would want to keep metadata in that group of key accesses on the same
server to reduce RPCs.

@@

% other use cases caches/tiers 
Storage systems have many other data management techniques that would benefit
from the caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. For example, Ceph administrators can use the
policies in ParSplice to automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain key access traces. We
hypothesize that since this is all software defined caching, something more
clever than LRU would improve cache utilization without sacrificing too much
performance.

>>>>

In the previous section, we used our data management language and the Mantle
policy engine to design effective cache management strategies for a new
application and storage system. In this section, we compare and contrast the
policies examined for file system metadata load balancing
in~\cite{sevilla:sc15-mantle} with the ones we designed and evaluated above for
cache management in ParSplice.

@@

The cache management policies we developed earlier can be used by load
balancing policies to effectively spread load across a cluster. For example,
distributed file systems that load balance file system metadata across a
dedicated metadata cluster could use the caching policies to determine what
metadata to move and when to move it.  To demonstrate this idea, we analyze a
3-day Lustre file system metadata trace, collected at LANL.  The trace is
anonymized so all file names are replaced with a unique identifier and we do
not know which applications are running. We visualize a 1 hour window of the
trace in Figure~\ref{fig:trace-atime}, where the dots are the file system
metadata reads in a 1 hour window.  The \(x\) axis is time and the \(y\) axis
is the file ID, listed in the order that file IDs appear in the trace.  The
groups of accesses look similar to the ParSplice key accesses in
Figure~\ref{fig:keyspace-zoomed}.

Although other access pattern detection algorithms are possible, we use the one
designed for cache management in Section~\S\ref{sec:regime-detection} with
slight modifications based on our knowledge of file systems\footnote{We
filtered out requests for key IDs less than 2000, as these are most likely path
traversal requests to higher parts of the namespace.}. The vertical lines in
Figure~\ref{fig:trace-atime} are the groups of accesses identified by the
algorithm; it successfully detects the largest group of key accesses that
starts at time 1000 seconds and ends at time 2200 seconds. File systems that
load balance file system metadata across a cluster would want to keep metadata
in that group of key accesses on the same server for locality and would want to
avoid migrating metadata to a different server until the group of key accesses
completes.

@@

% other use cases caches/tiers 
Storage systems have many other data management techniques that would benefit
from the caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. For example, Ceph administrators can use the
policies in ParSplice to automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain key access traces.

### 2017-11-27 paper/sections/6_conclusion.tex
paper: outline for #43
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Future Work}

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({\it e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Ultimately, we hope that this automation enables control of policies by
machines instead of administrators.

@@

Data management encompasses a wide range of techniques that vary by domain and
service. Yet, the techniques require policies that shape the decision making
and finding the best policies is a difficult, multi-dimensional problem.  We
iterate to a custom solution for our target application that uses workload
access patterns to size its caches. Without tuning or parameter sweeps, our
solution saves memory without sacrificing performance for a variety of initial
conditions, including the scale, duration, configuration, and hardware of the
simulation. More importantly, rather than attempting to construct a single,
complex policy that works for a variety of scenarios, we instead use the Mantle
framework to enable software-defined storage systems to flexibly change
policies as the workload changes.  We also observe that many of the primitives
and strategies have enough in common with data management in file systems that
they both can be expressed with similar semantics and a general policy engine.
Our prototype lays the groundwork for future work in adaptable policies and
policy generation, which ushers in an era where policies are controlled by
machines instead of administrators.

>>>>

@@

Data management encompasses a wide range of techniques that vary by application
and storage system.  Yet, the techniques require policies that shape the
decision making and finding the best policies is a difficult, multi-dimensional
problem.  We iterate to a custom solution for our target application that uses
workload access patterns to size its caches. Without tuning or parameter
sweeps, our solution saves memory without sacrificing performance for a variety
of initial conditions, including the scale, duration, configuration, and
hardware of the simulation. More importantly, rather than attempting to
construct a single, complex policy that works for a variety of scenarios, we
instead use the Mantle framework to enable software-defined storage systems to
flexibly change policies as the workload changes.  We also observe that many of
the primitives and strategies have enough in common with data management in
file systems that they both can be expressed with similar semantics and a
general policy engine.

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across
applications and storage systems.  The value of such a collection eases the
burden of policy development and paves the way for solutions that remove the
administrator from the development cycle, such as (1) adaptable policies that
automatically switch to new strategies when the current strategy behaves poorly
({\it e.g.}, thrashing, making no progress, etc.), and (2) policy generation,
where new policies are constructed automatically by examining the collection of
existing policies.  Ultimately, we hope that this automation enables control of
policies by machines instead of administrators.
